The evolution of gene expression in the term placenta of viviparous mammals

### 2016-04-18 manuscript/manuscript.Rnw
center and ragged common name
the second mention of horse is just \horse
document why we're using a ram

\begin{table}
  \begin{tabular}{c c c c}
    \toprule
    Species                       & Common Name & Gestation (days) & \multicolumn{1}{p{1in}}{\centering Average Litter Size}  & Source \\
    \midrule
    \textit{Ateles fusciceps}     & Spider monkey & 224 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Bos taurus}           & Cow & 278 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Canis familiaris}     & Dog & 61 & 4 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Dasypus novemcinctus} & Nine-banded Armadillo & 150 & 4 & \citet{Smith1984:amazingarmadillo} \\
    \textit{Equus caballus}       & Horse & 336 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation}\\
    \textit{Loxodonta africana}   & African savanna elephant & 644 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Homo sapiens}         & Human & 270 & 1.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Monodelphis domestica}& Gray short-tailed opossum & 15 & 7.1 & \citet{Harder.ea1993:Gestationandplacentationin} \\
    \textit{Mus musculus}         & Mouse & 20 & 6.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Spalax galili}   & Blind mole rat from Upper Galilee & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals} \\
    \textit{Ovis aries}           & Domestic sheep & 151 & 1.35 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Pan paniscus}         & Bonobo & 223 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental} \\
    \textit{Spalax carmeli}       & Blind mole rat from northern Israel & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals}\\
    \textit{Sus scrofa}           & Domestic pig (Duroc) & 114 & 5.8 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \bottomrule
  \end{tabular}
  \caption{Gestational periods in days and average liter sizes for studied species.}
  \label{tab:gestational_periods}
\end{table}

@@

\begin{figure}
<<species.tree,fig.width=8,fig.height=6,echo=FALSE>>=
species.tree <- 
    read.tree(text="(Monodelphis_domestica,((Dasypus_novemcinctus,Loxodonta_africana)Atlantogenata,((Canis_familiaris,(Equus_caballus,(Sus_scrofa,(Ovis_aries,Bos_taurus))))Laurasiatheria,(((Spalax_carmeli,Spalax_galili),Mus_Musculus)Rodentia,((Pan_paniscus,Homo_sapiens),Ateles_fuciceps)Primates)Euarchontoglires)Boreoeutheria)Eutheria);")
pp <- ggtree(species.tree)+
    geom_text(aes(label=label),hjust=0,size=5,fontface="italic")+
    scale_x_continuous(limits=c(0,10))
print(pp %>%
      ## monodelphis
      phylopic("91324e57-b3f1-42e0-abe3-43e5bc8aa4c6", color="#86B875",alpha=0.5,node=1,width=0.2)
      ## Armadillo
      %>% phylopic("5d59b5ce-c1dd-40f6-b295-8d2629b9775e", color="#44EEBB",alpha=0.5,node=2,width=0.15)
      ## elephant
      %>% phylopic("80db1004-bc9f-4318-84e7-cdd9639a1f3e",color="#BB1111",alpha=0.5,node=3,width=0.15)
      ## dog
      %>% phylopic("4d83a0cd-cf06-4a32-9a5a-0a6b644158c1",color="#FFAACC",alpha=0.5,node=4,width=0.15)
      ## horse
      %>% phylopic("071ee517-a0f1-4d19-aa29-812b9f86cb53",color="#22BBBB",alpha=0.5,node=5,width=0.15)
      ## pig
      %>% phylopic("008d6d88-d1be-470a-8c70-73625c3fb4fb",color="#33AA11",alpha=0.5,node=6,width=0.12)
      ## sheep
      %>% phylopic("cbea1964-89bf-4ce1-99b0-58fe0f9170b3",color="#004411",alpha=0.5,node=7,width=0.18)
      ## cow
      %>% phylopic("415714b4-859c-4d1c-9ce0-9e1081613df7",color="#881111",alpha=0.5,node=8,width=0.12)
      ## s carmeli
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#881188",alpha=0.5,node=9,width=0.1)
      ## n galili
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#888888",alpha=0.5,node=10,width=0.1)
      ## mouse
      %>% phylopic("6b2b98f6-f879-445f-9ac2-2c2563157025",color="#BB11CC",alpha=0.5,node=11,width=0.18)
      ## bonobo
      %>% phylopic("f06b4dd0-094e-4f37-9105-25fdb4eb1b02",color="#1111FF",alpha=0.5,node=12,width=0.18)
      ## spider monkey
      %>% phylopic("aceb287d-84cf-46f1-868c-4797c4ac54a8",color="#FF11FF",alpha=0.5,node=14,width=0.25)
      ## human
      %>% phylopic("a9f4ebd5-53d7-4de9-9e66-85ff6c2d513e",color="#CCCC44",alpha=0.5,node=13,width=0.18)
      )
@ 
\caption{Species tree for organisms studied showing the relationship
  between species. Silhouettes of species were obtained from
  \url{http://phylopic.org}. }
\label{fig:species_tree}
\end{figure}

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Equus caballus} are shown in
\cref{tab:top50_horse_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_equus_caballus_placenta}. The gene
with the highest expression was the relaxin,
\Sexpr{gene.expression("equus_caballus","RLN",complete=TRUE)},
followed by two miRNA
(\Sexpr{gene.expression("equus_caballus","ENSECAG00000025328",complete=TRUE)}
and
\Sexpr{gene.expression("equus_caballus","ENSECAG00000027056",complete=TRUE)})
with no known \textit{in silico} targets
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}, the thymosin
\Sexpr{gene.expression("equus_caballus","TMSB4",complete=TRUE)},
members of the insulin growth factor family
(\Sexpr{gene.expression("equus_caballus","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("equus_caballus","IGFBP3",complete=TRUE)}) and
multiple members of annexin complexes
(\Sexpr{gene.expression("equus_caballus","S100A6",complete=TRUE)},
\Sexpr{gene.expression("equus_caballus","S100A1",complete=TRUE)}, and
\Sexpr{gene.expression("equus_caballus","ANXA2",complete=TRUE)}),
similar to the pattern of expression seen in \pig and \dog.

>>>>

\begin{table}
  \begin{tabular}{c >{\raggedright \centering}p{1in} c c c}
    \toprule
    Species                       & Common Name & Gestation (days) & \multicolumn{1}{p{1in}}{\centering Average Litter Size}  & Source \\
    \midrule
    \textit{Ateles fusciceps}     & Spider monkey & 224 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Bos taurus}           & Cow & 278 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Canis familiaris}     & Dog & 61 & 4 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Dasypus novemcinctus} & Nine-banded Armadillo & 150 & 4 & \citet{Smith1984:amazingarmadillo} \\
    \textit{Equus caballus}       & Horse & 336 & 1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation}\\
    \textit{Loxodonta africana}   & African savanna elephant & 644 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental}\\
    \textit{Homo sapiens}         & Human & 270 & 1.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Monodelphis domestica}& Gray short-tailed opossum & 15 & 7.1 & \citet{Harder.ea1993:Gestationandplacentationin} \\
    \textit{Mus musculus}         & Mouse & 20 & 6.1 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Spalax galili}   & Blind mole rat from Upper Galilee & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals} \\
    \textit{Ovis aries}           & Domestic sheep & 151 & 1.35 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \textit{Pan paniscus}         & Bonobo & 223 & 1 & \citet{Ernest2003:lifehistorycharacteristicsofplacental} \\
    \textit{Spalax carmeli}       & Blind mole rat from northern Israel & 28-42 & 4 & \citet{Nevo1997:mosiacevolutionofsubterranianmammals}\\
    \textit{Sus scrofa}           & Domestic pig (Duroc) & 114 & 5.8 & \citet{Kiltie1982:intraspecificvariationmammaliangestation} \\
    \bottomrule
  \end{tabular}
  \caption{Gestational periods in days and average liter sizes for studied species.}
  \label{tab:gestational_periods}
\end{table}

@@

\begin{figure}
<<species.tree,fig.width=8,fig.height=6,echo=FALSE>>=
species.tree <- 
    read.tree(text="(Monodelphis_domestica,((Dasypus_novemcinctus,Loxodonta_africana)Atlantogenata,((Canis_familiaris,(Equus_caballus,(Sus_scrofa,(Ovis_aries,Bos_taurus))))Laurasiatheria,(((Spalax_carmeli,Spalax_galili),Mus_Musculus)Rodentia,((Pan_paniscus,Homo_sapiens),Ateles_fuciceps)Primates)Euarchontoglires)Boreoeutheria)Eutheria);")
pp <- ggtree(species.tree)+
    geom_text(aes(label=label),hjust=0,size=5,fontface="italic")+
    scale_x_continuous(limits=c(0,10))
print(pp %>%
      ## monodelphis
      phylopic("91324e57-b3f1-42e0-abe3-43e5bc8aa4c6", color="#86B875",alpha=0.5,node=1,width=0.2)
      ## Armadillo
      %>% phylopic("5d59b5ce-c1dd-40f6-b295-8d2629b9775e", color="#44EEBB",alpha=0.5,node=2,width=0.15)
      ## elephant
      %>% phylopic("80db1004-bc9f-4318-84e7-cdd9639a1f3e",color="#BB1111",alpha=0.5,node=3,width=0.15)
      ## dog
      %>% phylopic("4d83a0cd-cf06-4a32-9a5a-0a6b644158c1",color="#FFAACC",alpha=0.5,node=4,width=0.15)
      ## horse
      %>% phylopic("071ee517-a0f1-4d19-aa29-812b9f86cb53",color="#22BBBB",alpha=0.5,node=5,width=0.15)
      ## pig
      %>% phylopic("008d6d88-d1be-470a-8c70-73625c3fb4fb",color="#33AA11",alpha=0.5,node=6,width=0.12)
      ## sheep -- this is a male, the sample is female, but there aren't any females in phylopic
      %>% phylopic("cbea1964-89bf-4ce1-99b0-58fe0f9170b3",color="#004411",alpha=0.5,node=7,width=0.18)
      ## cow
      %>% phylopic("415714b4-859c-4d1c-9ce0-9e1081613df7",color="#881111",alpha=0.5,node=8,width=0.12)
      ## s carmeli
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#881188",alpha=0.5,node=9,width=0.1)
      ## n galili
      %>% phylopic("287c58ad-009b-4878-9a9e-59fd7e0fe66f",color="#888888",alpha=0.5,node=10,width=0.1)
      ## mouse
      %>% phylopic("6b2b98f6-f879-445f-9ac2-2c2563157025",color="#BB11CC",alpha=0.5,node=11,width=0.18)
      ## bonobo
      %>% phylopic("f06b4dd0-094e-4f37-9105-25fdb4eb1b02",color="#1111FF",alpha=0.5,node=12,width=0.18)
      ## spider monkey
      %>% phylopic("aceb287d-84cf-46f1-868c-4797c4ac54a8",color="#FF11FF",alpha=0.5,node=14,width=0.25)
      ## human
      %>% phylopic("a9f4ebd5-53d7-4de9-9e66-85ff6c2d513e",color="#CCCC44",alpha=0.5,node=13,width=0.18)
      )
@ 
\caption{Species tree for organisms studied showing the relationship
  between species. Silhouettes of species were obtained from
  \url{http://phylopic.org}. }
\label{fig:species_tree}
\end{figure}

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \horse are shown in
\cref{tab:top50_horse_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_equus_caballus_placenta}. The gene
with the highest expression was the relaxin,
\Sexpr{gene.expression("equus_caballus","RLN",complete=TRUE)},
followed by two miRNA
(\Sexpr{gene.expression("equus_caballus","ENSECAG00000025328",complete=TRUE)}
and
\Sexpr{gene.expression("equus_caballus","ENSECAG00000027056",complete=TRUE)})
with no known \textit{in silico} targets
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}, the thymosin
\Sexpr{gene.expression("equus_caballus","TMSB4",complete=TRUE)},
members of the insulin growth factor family
(\Sexpr{gene.expression("equus_caballus","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("equus_caballus","IGFBP3",complete=TRUE)}) and
multiple members of annexin complexes
(\Sexpr{gene.expression("equus_caballus","S100A6",complete=TRUE)},
\Sexpr{gene.expression("equus_caballus","S100A1",complete=TRUE)}, and
\Sexpr{gene.expression("equus_caballus","ANXA2",complete=TRUE)}),
similar to the pattern of expression seen in \pig and \dog.
